---
title: "Exploratory Data"
author: "Gauchet van Antwerpen"
date: "4/29/2020"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
#rm(list = ls());
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_libraries, message=FALSE, , echo=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
library(tidyverse) # handy utility functions
library(outliers) # library for identifying outliers
library(forecast)
library(ggpubr)
```

```{r load_local_vars, message=FALSE, , echo=FALSE}

fp_local_data <- "/Users/gva/Documents/02 Werk/00 Appels en Pere/02 Projekte/02 ACCTech/09 data"


#my.dat.mining.all <- readRDS(file.path(fp_local_data, "gangs_all.rds"));

my.dat.gangs.over_abs <- readRDS(file.path(fp_local_data, "gangs_abs_over.rds"));

#saveRDS(my.dat.gangs.over_abs, file.path(fp_local_data, "gangs_abs_over.rds"));



#my.dat.gangs.over_abs$my_abs$GNG_NAME
#my.dat.gangs.over_abs$my_over$GNG_NAME <- my.dat.gangs.over_abs$my_over$XT_GangName;

#my.dat.gangs.over_abs$my_over$XT_GangName <- NULL;

#my.dat.gangs.over_abs$my_over$DateKey <- as.Date(my.dat.gangs.over_abs$my_over$DateKey, format = "%Y-%m-%d");

#my.dat.gangs.over_abs$my_abs$DateKey <- as.Date(my.dat.gangs.over_abs$my_abs$DateKey, format = "%Y/%m/%d");

#my.dat.gangs.over_abs$my_over$GNG_NAME <- as.factor(my.dat.gangs.over_abs$my_over$GNG_NAME);

#my.dat.gangs.over_abs$my_abs$GNG_NAME <- as.factor(my.dat.gangs.over_abs$my_abs$GNG_NAME);

#my.dat.gangs.over_abs$my_abs$n_abs <- my.dat.gangs.over_abs$my_abs$NumberOfAbsences;

#my.dat.gangs.over_abs$my_abs$NumberOfAbsences <- NULL;

#my.dat.gangs.over_abs$my_abs$n_u_abs <- my.dat.gangs.over_abs$my_abs$NumberOfUnplannedAbsences;

#my.dat.gangs.over_abs$my_abs$NumberOfUnplannedAbsences <- NULL;

#my.dat.gangs.over_abs$my_over$n_ovr_hrs <- my.dat.gangs.over_abs$my_over$TotalOvertimePayHours;

#my.dat.gangs.over_abs$my_over$TotalOvertimePayHours <- NULL;

```

# General Stats

```{r Outliers_calc, echo=FALSE}
#my_gangs <- dat_gangs

#my_mining <- dat_mining

# get the z-scores for each value in absence, overtime is already focussed
#outlier_scr_n_over <- scores(my_gangs$n_over)
outlr_scr_n_abs <- scores(my.dat.gangs.over_abs$my_abs$n_abs);

outlr_scr_n_u_abs <- scores(my.dat.gangs.over_abs$my_abs$n_u_abs);

# create a logical vector the same length as outlier_scores
# that is "TRUE" if outlier_scores is greater than 3 or
# less than negative 3
is_outl_n_abs <- outlr_scr_n_abs > 3 | outlr_scr_n_abs < -3;
is_outl_n_u_abs <- outlr_scr_n_u_abs > 3 | outlr_scr_n_u_abs < -3;


# add a column with info whether the overtime is an outlier
my.dat.gangs.outl_n_abs <- is_outl_n_abs;
my.dat.gangs.outl_n_u_abs <- is_outl_n_u_abs;




# create a dataframe with only outliers
my.dat.gangs.abs.outl <- my.dat.gangs.over_abs$my_abs[outlr_scr_n_abs > 3| outlr_scr_n_abs < -3, ];

my.dat.gangs.u_abs.outl <- my.dat.gangs.over_abs$my_abs[outlr_scr_n_u_abs > 3| outlr_scr_n_u_abs < -3, ];

#gang_abs_outl_01$is_outl_n_abs <- NULL
#gang_abs_outl_01$is_outl_n_ovr <- NULL
#gang_abs_outl_01$G1 <- NULL
#gang_abs_outl_01$G2 <- NULL
#gang_abs_outl_01$G3 <- NULL
#gang_abs_outl_01$G4 <- NULL

#outlier_scr_n_abs_2 <- scores(gang_abs_outl_01$n_abs)
#is_outl_n_abs_2 <- outlier_scr_n_abs_2 > 3 | outlier_scr_n_abs_2 < -3
#my_gangs_2 <- is_outl_n_abs_2
#my_gangs_2$is_outl_n_abs_2 <- is_outl_n_abs_2

#gang_abs_outl_02 <- gang_abs_outl_01[outlier_scr_n_abs_2 > 3| outlier_scr_n_abs_2 < -3, ]

#gang_abs_outl_02$is_outl_n_abs_2 <- is_outl_n_abs_2

#print("Outliers scores are satistically determined via a Z-score.")
```

Outliers for Mining Absentiesm are satistically determined via the statiscally Z-score.


## 1 Overall summary of the COMPLETE Absenteism mining data set:

```{r Summary Absenteism, echo=FALSE, warning=FALSE}
#print("1 Overall summary of the COMPLETE mining data set:");
#gang_abs_outl_01
#gang_over_outl_01
#my_gang_grp
#gang_abs_outl_01$n_over <- NULL
####################################Absence COMPLETE
print(paste("2 Summary of Absence COMPLETE, data set size: ", length(my.dat.gangs.over_abs$my_abs$GNG_NAME))); 

#print("Stats summary for all absence")

summary(my.dat.gangs.over_abs$my_abs);

```

From the information above the following are observed:

- High absenteism days are
-- Admin Guesthouse
-- Crusher Gangs 1-4
-- Executive

- Average of 1.451 days planned and 0.793 unplanned

- Max days absenteism 90.000 days taken and 83.000 unplanned

## 2 Overall summary of the Focussed Absenteism mining data set:

```{r Summary Focussed Absenteism, echo=FALSE, warning=FALSE}
####################################Absence GANGS
my.dat.foc.abs <- semi_join(my.dat.gangs.over_abs$my_abs, my.dat.gangs.over_abs$my_over, by = "GNG_NAME");

print(paste("Summary of focussed group absence, data set size: ", length(my.dat.foc.abs$GNG_NAME)));

#print("Stats summary for focussed absence")

summary(my.dat.foc.abs);


```

From the information above the following are observed:

- Average of 7.582 days planned and 4.142 unplanned

- Max days absenteism 41.000 days taken and 33.000 unplanned

## 3 Overall summary of the Focussed Overtime mining data set:
```{r Focussed Overtime, echo=FALSE, warning=FALSE}

print(paste("Summary of overtime mining, data set size: ", length(my.dat.gangs.over_abs$my_over$GNG_NAME)));

summary(my.dat.gangs.over_abs$my_over);
```

From the information above the following are observed:

- Average of 58.43 hours taken

- Max days overtime 1042.00 hours

# Histograms

## Absenteism Histograms
```{r Histograms Abs, echo=FALSE , warning=FALSE}

# Histograms for abs

print("Absenteism All of mine...")

ggplot(data = my.dat.gangs.over_abs$my_abs, aes(x = n_abs)) +
geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 6) +
geom_density(alpha=.2, fill="blue") +
labs(x = 'Absenteism', y = 'Density')



# Histograms for absence
#print("Absenteism Histogram")

#print("Absenteism: All of mine")

#ggplot(data = my_mining, aes(x = n_abs/1e6)) +
#  geom_histogram(aes(y=..density..), colour="black", fill="white", bins = #12) +
#  geom_density(alpha=.2, fill="green") +
#  labs(x = 'Remove', y = 'Density')


#print("Absenteism: Outliers only level")

#ggplot(data = gang_abs_outl_02, aes(x = n_abs/1e6)) +
#  geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 12) +
#  geom_density(alpha=.2, fill="green") +
#  labs(x = '', y = 'Density')



print("Histograms are highly skewed to the left which indicates outliers. Outliers is an indication of something outside the mean. Mean in this case is average hourse");

```

## Absenteism: Focussed Histograms

```{r Histograms Abs Focussed, echo=FALSE}

#print("Absenteism: Focussed")

ggplot(data = my.dat.foc.abs, aes(x =n_abs)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 12) +
  geom_density(alpha=.2, fill="green") +
  labs(x = 'Focussed Group', y = 'Density');
```

## Overtime: Focussed Histograms
```{r Histograms ovr Focussed, echo=FALSE}

#print("Overtime Focussed group: ")

ggplot(data = my.dat.gangs.over_abs$my_over, aes(x = n_ovr_hrs)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 12) +
  geom_density(alpha=.2, fill="orange") +
  labs(x = 'Overtime Focussed', y = 'Density')

```

```{r outliers not used, echo=FALSE}
#my_gangs <- dat_gangs

#my_mining <- dat_mining

# use a box plot to see if we have outliers in the overtime column
# of the my_gangs dataset
#ggplot(my_gangs, aes(x = "overtime", y = n_over)) +
#  geom_boxplot()

# use a box plot to see if we have outliers in the overtime column
# of the data_v2 dataset
#ggplot(my_gangs, aes(x = "absence", y = n_abs)) +
#  geom_boxplot()

# look for outliers for each different GNG_NAME description
#ggplot(my_gangs, aes(GNG_NAME, n_over)) +
#  geom_boxplot() +
#  coord_flip() # flip our plot so the labels are easier to read

# look for outliers for each different GNG_NAME description
#ggplot(my_gangs, aes(GNG_NAME, n_abs)) +
#  geom_boxplot() +
#  coord_flip() # flip our plot so the labels are easier to read

```

# Correlations

```{r correlations, echo=FALSE, warning= FALSE}
my.dat.foc.all <- inner_join(my.dat.gangs.over_abs$my_over,my.dat.foc.abs);


my_corr_foc <- cor(my.dat.foc.all[,3:4])

print("Standard corrplot")

p <- ggplot(my.dat.foc.all, aes(x=n_ovr_hrs, y=n_abs)) +
  geom_point(aes(col=my.dat.foc.all$GNG_NAME)) +
  geom_smooth(method = lm) +
  stat_cor(method = "pearson", label.x = 20)
p

print("method= loess");


gg <- ggplot(my.dat.foc.all, aes(x=n_ovr_hrs, y=n_abs)) + 
  geom_point(aes(col=my.dat.foc.all$GNG_NAME)) + 
  geom_smooth(method="loess", se=F) + 
  #xlim(c(0, 0.1)) + 
  #ylim(c(0, 500000)) + 
  labs(subtitle="Overtime Vs Absense", 
       y="Absence days", 
       x="Overtime Hours", 
       title="Scatterplot");
       #caption = "Source: midwest")

plot(gg)
#my.dat.foc.abs

#grob_foc_abs <- grobTree(textGrob(paste("Pearson Correlation : ", round(cor(my.dat.foc.abs[,3:4]), 4)) ), gp = gpar(col = "red", fontsize = 11, fontface = "bold")); #x = 0.63, y = 0.97, hjust = 0, 

#plt_grob_foc_abs <- ggplot(my.dat.foc.abs, aes(x=n_abs, y=n_u_abs)) + geom_point() + ggtitle("Absense vs Unplanned") + geom_smooth(method=lm, se=FALSE) + scale_x_continuous(name = "n_abs", limits = c(5, 15), breaks = seq(5, 15, 2)) + scale_y_continuous(name = "n_u_abs", limits = c(5,15), breaks = seq(5, 15, 2)) + annotation_custom(grob1) + theme(plot.title = element_text(hjust = 0.5))





#my.dat.foc.all$n_ovr_hrs

#grob_foc_all <- grobTree(textGrob(paste("Pearson Correlation : ", round(cor(my.dat.foc.all[,3:4]), 4)) ), gp = gpar(col = "red", fontsize = 11, fontface = "bold")); #x = 0.63, y = 0.97, hjust = 0, 

#plt_grob_foc_over <- ggplot(my.dat.foc.all, aes(x=n_ovr_hrs, y=n_abs)) + geom_point() + ggtitle("Absense vs Unplanned") + geom_smooth(method=lm, se=FALSE) + scale_x_continuous(name = "Overtime Hours", limits = c(5, 15), breaks = seq(5, 15, 2)) + scale_y_continuous(name = "Absense days", limits = c(5,15), breaks = seq(5, 15, 2)) + annotation_custom(grob1) + theme(plot.title = element_text(hjust = 0.5))

```

The following can be noted from the above correlation plots:
- Correlation coffiecent R is defined (0 < R < 1)
- R for overtime and absence is 0.36 wich indicate that there is a correlation
- p stat < 2.2*e^-16


# Showing Box Plots

## Absence COMPLETE data set

```{r outl_box_abs, echo=FALSE, warning=FALSE}
#print("Detail of the COMPLETE data set: ");

# look at the boxlpot of the complete mining
#my.dat.gangs.over_abs$my_abs$abs_outl <- my.dat.gangs.abs.outl
#my.dat.gangs.outl_n_abs
#my.dat.gangs.outl_n_u_abs
#print("Complete mining set")
ggplot(my.dat.gangs.over_abs$my_abs, aes(GNG_NAME, n_abs)) +
  geom_boxplot();# +
  coord_flip();# +
  #facet_wrap(~my.dat.gangs.outl_n_abs);


print("Outlier mining set")

ggplot(my.dat.gangs.abs.outl, aes(GNG_NAME, n_abs)) +
  geom_boxplot();# +
  coord_flip();

print("Outlier focussed set")
# look at the boxplot for overtime outliers
ggplot(my.dat.foc.abs, aes(GNG_NAME, n_abs)) +
  geom_boxplot()# +
#  coord_flip() # flip our plot so the labels are easier to read


```

## Overtime FOCUSSED data set

```{r foc_box, echo=FALSE}
#print("Detail of the OUTLIER mining data set for Overtime:");



# look at the boxplot for absence level one
ggplot(my.dat.gangs.over_abs$my_over, aes(GNG_NAME, n_ovr_hrs)) +
  geom_boxplot() +
  coord_flip() # flip our plot so the labels are easier to read

# look at the boxplot for absence level one
#gang_abs_outl_02$n_abs <- sort(gang_abs_outl_02$n_abs, decreasing = TRUE)
#ggplot(gang_abs_outl_02, aes(GNG_NAME, n_abs)) +
#  geom_boxplot() +
#  coord_flip() # flip our plot so the labels are easier to read
```


# Dive into Seasonality

This section is based on https://towardsdatascience.com/forecasting-with-r-trends-and-seasonality-def24280e71f

```{r seasonality, echo=FALSE}

#library(tstools)
#install.packages("tstools")
#install.packages("devtools")
#library(devtools)

#my_gng_abs_out_ml



#tss_abs <- ts(my.dat.gangs.over_abs$my_abs[,3], start = c(2018,1), end  = c(2020,1), frequency = 24);
#plot(tss_abs)#ylim= c(1,1100)

#tss_over <- ts(my.dat.gangs.over_abs$my_over[,3], start = c(2018,1), end  = c(2020,1), frequency = 24);
#plot(tss_over)#, ylim= c(1,1100)

#seasonplot(tss_abs)

#seasonplot(tss_over)

```

## Decomposition

```{r decomp, echo=FALSE}

#print("Absence seasonality")


my_abs_piv <- my.dat.gangs.over_abs$my_abs %>%
  pivot_longer(cols = starts_with(c("n_")), names_to = 'Gangs', values_to = 'Value') %>%
  group_by(DateKey) %>%
  summarise(ValueMean = mean(as.numeric(Value)))

# range(data$Date2)
my_abs_piv_ts <- ts(data = my_abs_piv$ValueMean, frequency = 365, start = c(2018,1), end  = c(2020,1))

my_decom_abs <- decompose(my_abs_piv_ts)

#my_decom$trend

p_seasonality_abs <- plot(my_decom_abs$trend)

#print("Overtime seasonality Focussed")

my_over_piv <- my.dat.gangs.over_abs$my_over %>%
  pivot_longer(cols = starts_with(c("n_")), names_to = 'Gangs', values_to = 'Value') %>%
  group_by(DateKey) %>%
  summarise(ValueMean = mean(as.numeric(Value)))

# range(data$Date2)
my_over_piv_ts <- ts(data = my_over_piv$ValueMean, frequency = 365, start = c(2018,1), end  = c(2020,1));

my_decom_over <- decompose(my_over_piv_ts)

p_seasonality_over <- plot(my_decom_over$trend)

#multiplot(p_seasonality_abs, p_seasonality_over)

```

From the information above it can be noted the trends of overtime vs absence are similar


```{r multiplot, echo=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```


```{r template2, echo=FALSE}


```
